---
title: "Naive Bayes Demo - Project 1"
author: "Niko Montez Graham"
date: "2024-05-31"
output: pdf_document
---
## Importing Libraries and Setting the Seed
```{r}
library(caTools)
library(e1071)
set.seed(42)
```

## Reading and Cleaning Data 
```{r}
NBdata <- read.csv('/Users/nikomg24/NBcustomer_data.csv')
summary(NBdata)
```
```{r}
head(NBdata, 5)
```

- Now that we have an idea of what kind of data we are looking into, lets make sure the data is clean and free form null values: 
```{r}
colSums(is.na(NBdata))
```
- Luckily, no instances need to be dropped in this case! Now, we can move on to the next phase of the Naive Bayes method. 

## Displaying Target Varible as Categorical 
```{r}
NBdata$Purchase <- as.factor(NBdata$Purchase)
summary(NBdata$Purchase)
```
- Now that the target variable is in this form, we can now begin the next step in this process!

## Splitting the Data
```{r}
mask = sample.split(NBdata$Purchase, SplitRatio = 0.80)
str(mask)
```
```{r}
traindata <- subset(NBdata, mask == TRUE)
testdata <- subset(NBdata, mask == FALSE)
```
- Perfect! Now that we have our train and testing data, we can move on to the final few steps of the process!

## Creating the Naive Bayes Model
- In my project, I will be creating a few different models in order to see which variables lead to more accurate predictions of the classifier. Let's create those now:
```{r}
model1 <- naiveBayes(x = traindata[,c('Visits','Product_Views')], y = traindata$Purchase)
model2 <- naiveBayes(x = traindata[,c('Age','Income')], y = traindata$Purchase)
model3 <- naiveBayes(x = traindata[,c('Income','Product_Views')], y = traindata$Purchase)
ypred1 <- predict(model1, newdata = testdata[,c('Visits','Product_Views')])
ypred2 <- predict(model2, newdata = testdata[,c('Age','Income')])
ypred3 <- predict(model3, newdata = testdata[,c('Income','Product_Views')])
```

## Plotting the Test Data for Each Model 
```{r}
data = testdata

minX1 = min(data[,4]); maxX1 = max(data[,4]); range1 = diff(range(data[,4]))
minX2 = min(data[,6]); maxX2 = max(data[,6]); range2 = diff(range(data[,6]))
len = 120
X1 = seq(from=minX1-0.1*range1, to=maxX1+0.1*range1, length.out=len)
X2 = seq(from=minX2-0.1*range2, to=maxX2+0.1*range2, length.out=len)

grid_data = expand.grid(X1, X2)  
colnames(grid_data) = c("Visits", "Product_Views")

y_grid = predict(model1, newdata = grid_data)
par(mar=c(4,4,1,4)) 
plot(data[,c("Visits", "Product_Views")],
     main = "Purchase Classification",
     xlab = "Visits", ylab = "Product Views",
     xlim = range(X1), ylim = range(X2))

points(grid_data, pch = '.', col = rgb(0,0,0,alpha=0.5))
points(data[data$Purchase == "Yes", c("Visits", "Product_Views")], pch = 21, bg = "green4")
points(data[data$Purchase != "Yes", c("Visits", "Product_Views")], pch = 21, bg = "red3")
par(mar=c(4,4,1,4)) 
plot(data[,c('Visits','Product_Views')],
     main = "Purchase Classification",
     xlab = "Visits", ylab = "Product Views",
     xlim = range(X1), ylim = range(X2))
contour(X1, X2, matrix(as.numeric(y_grid), length(X1), length(X2)), add = TRUE)
points(grid_data[y_grid == "Yes",], pch = '.', col = "green3")
points(grid_data[y_grid != "Yes",], pch = '.', col = "red")
points(data[data$Purchase == "Yes", c('Visits','Product_Views')], pch = 21, bg = "green4")
points(data[data$Purchase != "Yes", c('Visits','Product_Views')], pch = 21, bg = "red3")
```
```{r}
data = testdata

minX1 = min(data[,1]); maxX1 = max(data[,1]); range1 = diff(range(data[,1]))
minX2 = min(data[,3]); maxX2 = max(data[,3]); range2 = diff(range(data[,3]))
len = 120
X1 = seq(from=minX1-0.1*range1, to=maxX1+0.1*range1, length.out=len)
X2 = seq(from=minX2-0.1*range2, to=maxX2+0.1*range2, length.out=len)

grid_data = expand.grid(X1, X2)  
colnames(grid_data) = c('Age','Income')

y_grid = predict(model2, newdata = grid_data)
par(mar=c(4,4,1,4)) 
plot(data[,c('Age','Income')],
     main = "Purchase Classification",
     xlab = "Age", ylab = "Income",
     xlim = range(X1), ylim = range(X2))

points(grid_data, pch = '.', col = rgb(0,0,0,alpha=0.5))
points(data[data$Purchase == "Yes", c('Age','Income')], pch = 21, bg = "green4")
points(data[data$Purchase != "Yes", c('Age','Income')], pch = 21, bg = "red3")
par(mar=c(4,4,1,4)) 
plot(data[,c('Age','Income')],
     main = "Purchase Classification",
     xlab = "Age", ylab = "Income",
     xlim = range(X1), ylim = range(X2))
contour(X1, X2, matrix(as.numeric(y_grid), length(X1), length(X2)), add = TRUE)
points(grid_data[y_grid == "Yes",], pch = '.', col = "green3")
points(grid_data[y_grid != "Yes",], pch = '.', col = "red")
points(data[data$Purchase == "Yes", c('Age','Income')], pch = 21, bg = "green4")
points(data[data$Purchase != "Yes", c('Age','Income')], pch = 21, bg = "red3")
```
```{r}
data = testdata

minX1 = min(data[,3]); maxX1 = max(data[,3]); range1 = diff(range(data[,3]))
minX2 = min(data[,6]); maxX2 = max(data[,6]); range2 = diff(range(data[,6]))
len = 120
X1 = seq(from=minX1-0.1*range1, to=maxX1+0.1*range1, length.out=len)
X2 = seq(from=minX2-0.1*range2, to=maxX2+0.1*range2, length.out=len)

grid_data = expand.grid(X1, X2)  
colnames(grid_data) = c('Income','Product_Views')

y_grid = predict(model3, newdata = grid_data)
par(mar=c(4,4,1,4)) 
plot(data[,c('Income','Product_Views')],
     main = "NBdata (Training Data)",
     xlab = "Income", ylab = "Product Views",
     xlim = range(X1), ylim = range(X2))

points(grid_data, pch = '.', col = rgb(0,0,0,alpha=0.5))
points(data[data$Purchase == "Yes", c('Income','Product_Views')], pch = 21, bg = "green4")
points(data[data$Purchase != "Yes", c('Income','Product_Views')], pch = 21, bg = "red3")
par(mar=c(4,4,1,4)) 
plot(data[,c('Income','Product_Views')],
     main = "Purchase Classification",
     xlab = "Income", ylab = "Product Views",
     xlim = range(X1), ylim = range(X2))
contour(X1, X2, matrix(as.numeric(y_grid), length(X1), length(X2)), add = TRUE)
points(grid_data[y_grid == "Yes",], pch = '.', col = "green3")
points(grid_data[y_grid != "Yes",], pch = '.', col = "red")
points(data[data$Purchase == "Yes", c('Income','Product_Views')], pch = 21, bg = "green4")
points(data[data$Purchase != "Yes", c('Income','Product_Views')], pch = 21, bg = "red3")
```


## Confusion Matrices
```{r}
confM1 = table(testdata[ ,"Purchase"], ypred1)
confM2 = table(testdata[ ,"Purchase"], ypred2)
confM3 = table(testdata[ ,"Purchase"], ypred3)
confM1
confM2
confM3
```
As we can see from the matrices, the third model produced the most accurate predictions, only improperly predicting 6/40. This is information that could be very helpful for the business when trying to ensure a purchase be made more often by customers that use the site. 
